Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-20 ba_resubmit/kundinger_resubmit.tex
update.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

@@

\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.

In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

@@

\textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}

@@

>>>>

\textcolor{blue}{In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.}

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs and (2) compute the pairwise record comparisons over large datasets via parallel computing. \textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
 \textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking. This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking. We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}

%We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

@@

\textcolor{red}{(I don't love this right here, but I don't really know where else to put it. But if it won't cause a problem with the reviewers, I'll just leave it.)}
\textcolor{blue}{Perhaps any parts that aren't needed can be removed. I'll take a stab at it.}

%In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeler chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
%
%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

@@

\textcolor{red}{I actually don't like the $\Phi$ notation. It seems unnecessary. The only reason it's here is because Sadinle used it. What do you guys think?}
\textcolor{blue}{I would leave it since they didn't complain unless you have a better idea.}

@@

\textcolor{blue}{This needs to address the concern of R3, where we didn't cite Sadinle/FS. I would also suggest that we take the time to elaborate on some future work.}
